---
title: "Problem Set 2"
author: 3032247297
date: 2/10/2019
output: html_notebook
---


#Reading Data into R
```{r}
loyn <- read.csv("loyn.csv", header = TRUE)
#QUestion A
head(loyn, 5)

```

#Simple data curation
```{r}
loyn <- loyn[, c("ABUND", "ALT", "YR.ISOL", "L10AREA", "L10DIST", "L10LDIST", "GRAZE")]
loyn[["YEARS"]] <- 1983 - loyn[["YR.ISOL"]] ## 1983 reproduces the Box 6.2 result
```


#Multiple Regression
```{r}
fit_graze_years <- lm(ABUND ~ (GRAZE + YEARS), data = loyn)
## Notice how the "data" argument is one way to tell the lm function
## what data frame has the ABUND, GRAZE and YEARS vector


summary(fit_graze_years)
```


##QUestion B
```{r}
#1) y = 34.38 - 4.44x - 0.05x
#2)
confint(fit_graze_years, "YEARS", level = 0.95)
#3)
sum(resid(fit_graze_years)**2)
#4) The null hypothesis is that there is no association between the years variable and the abundance variable
#5)The null hypothesis for this variable is that there is no association between any of the variables. In this case that graze and years have an affect on abundance.
```

#Comparing reduced and full models
##Question C
1)
```{r}
Full <- lm(ABUND ~  L10AREA + GRAZE + L10DIST + YEARS, data = loyn)
summary(Full)
```

```{r}
reduced <- lm(ABUND ~ L10AREA + GRAZE, data = loyn)
summary(reduced)
```

2)
```{r}
residred <- sum(resid(reduced)**2)
residred
```

3)
```{r}
residfull <- sum(resid(Full)**2)
residfull
```

4)
```{r}
residred - residfull
```

5)
```{r}
(residred - residfull)/2
```

6)
```{r}
F <- anova(Full, reduced)["2", "F"]
F
```

7)
```{r}
numeratordiff <- reduced$df.residual - Full$df.residual
numeratordiff

denominatordiff = Full$df.residual
denominatordiff

```

8)
When conducting an anova test the F stat tells you if the difference in means between the two populations are statistically significant. Since F != 1, we reject the null

9)
```{r}
summary(Full)

4.849e-12
```
10)
We reject the null because our p value is way below our threshold.

11)
More variables are being used for the full model so the standard error is larger because of that. In addition the error is inversely proportional to the sample size.

12)
```{r}
estimate <- -0.6239 * -1
error <- 0.0431
sd <- estimate/error
per <- pt(sd, 48) * 100
per
```

13)
```{r}
100 - per
```

14)
The p value for fit 2 is smaller so we reject the null, and since it is large for fit 4 that means that we fail to reject the null.

#Question D
```{r}
simulate_d <- function(L10AREA, GRAZE, true_intercept, L10AREA_true_slope, GRAZE_true_slope, residual_std, n) {
  ABUND <- numeric(n)
  for (i in 1:n){
    ABUND[i] = true_intercept + (L10AREA[i] * L10AREA_true_slope)* (GRAZE[i] * GRAZE_true_slope) + rnorm(1, 0 ,residual_std)
  }
  return(data.frame(ABUND, L10AREA, GRAZE))
}
```


#Question E
```{r}
num_simulations <- 1000
GRAZE_p_values <- numeric(num_simulations)
GRAZE_slopes <- numeric(num_simulations)
set.seed(12345)
for(i in 1:num_simulations) {
  sim_data <- simulate_d(loyn[["L10AREA"]], loyn[["GRAZE"]], 22, 6.5, -1.0, 6.0, 56)
  fit <- lm(ABUND ~ L10AREA + GRAZE, data = sim_data)
  GRAZE_p_values[i] <- summary(fit)$coefficients['GRAZE', 'Pr(>|t|)']
  GRAZE_slopes[i] <- summary(fit)$coefficients['GRAZE', 'Estimate']
}
# Now you can look at the distribution of p-values and estimated slopes:
hist(GRAZE_p_values, main = "P-values for GRAZE")
hist(GRAZE_slopes, main = "Estimated coefficients for GRAZE")
# Finally, you can look at the statistical power: the probability of rejecting the null 
# when it is FALSE. From simulations, this is approximately the frequency of rejecting 
# the null.  We can calculate that as follows:
sum(GRAZE_p_values < 0.05) / num_simulations
# or, equivalently:
mean(GRAZE_p_values < 0.05)
```

1)
Just by looking at the graphs it seems like the estimated slope is not close to the true slope for graze

2)
The distribution would decrease since the estimated coefficients are closer to -2 than -1

























