---
title: "problem set 10"
author: "3032247297"
output:
  html_document:
    df_print: paged
---

I collaborated with nobody on this assignment.

Question A


```{r}
RK <- read.table("RoadKills.txt", header = TRUE)
summary(RK)
with(RK, plot(D.PARK, TOT.N, pch = 19, main = "Road Kill Data from Zuur et al. (2009)"))
```


Question B


```{r}
fit1 <- glm(TOT.N ~ D.PARK, family = poisson, data = RK)
summary(fit1)
logLik(fit1)
AIC(fit1)
deviance(fit1)
{
  with(RK, plot(D.PARK, TOT.N, pch = 19, main = "Road Kill Data from Zuur et al. (2009)"))
  with(RK, points(D.PARK, fitted(fit1), pch = 'x', col = 'red')) 
}

plot(RK$D.PARK, resid(fit1))

```

Question C

I think the first violation is that the errors are not independent, and the second is that the data does not seem to have constant variance. As X increases, the data gets clustered closer and closer together which indicates a decrease in variance and std deviation.It also does not look like there is a linear relationship between the response and the linear predictor, it seems almost sinusoidal, this is another violation of the GLM assumption.

Question D

```{r}
RK$D.PARK[15]
RK$TOT.N[15]

RK
```

1) 15(7187.762) + 33 = 107,849
2) 4.316 - 15(.000106) = 4.314
3)
```{r}
summary(RK[15])

u <- 1.5617

exp(-u)*(u^4.314)/factorial(4.314)
```
The answer is .0369

Question E

```{r}
logLikelihood <- function(intercept, slope) {
  sum(dpois(RK$TOT.N, exp(intercept + slope*RK$D.PARK), log = TRUE))
}
logLikelihood(4.316e+00, -1.059e-04)

slope.grid <- -0.000106 + seq(-9e-6, 9e-6, length = 21)
cbind(slope.grid) ## cbind shows the values vertically, which is easier to read.
intercept.grid <- 4.32 + seq(-9e-2, 9e-2, length = 21)
cbind(intercept.grid)
z <- matrix(0, nrow = length(intercept.grid), ncol = length(slope.grid))
for(i in 1:length(intercept.grid))
  for(j in 1:length(slope.grid))
    z[i, j] <- logLikelihood(intercept.grid[i], slope.grid[j])
z
{
  contour(intercept.grid, slope.grid, z, 
          levels = max(z) - (1:10), main = "log likelihood contours")
  points(4.32, -0.000106, pch = 'x', col = 'red')
}
```

The log likelihood surface looks like tree rings comprised of ellipses.

```{r}
fit2 <- glm(TOT.N ~ D.PARK, family = quasipoisson, data = RK)
summary(fit2)
```

Question F

The reported dispersion parameter is 7.630

Question G

The parameter estimates are nearly the same while the standard error is bigger and the resulting confidence intervals are also larger. The p value for Dpark is also less significant in fit2 than in fit1 but still very significant.

```{r}
logLik(fit2)
AIC(fit2)
deviance(fit2)
```

Question H

The deviance is the same in fit2 as fit1.

```{r}
RK$SQ.POLIC<-sqrt(RK$POLIC)
RK$SQ.WATRES<-sqrt(RK$WAT.RES)
RK$SQ.URBAN<-sqrt(RK$URBAN)
RK$SQ.OLIVE<-sqrt(RK$OLIVE)
RK$SQ.LPROAD<-sqrt(RK$L.P.ROAD)
RK$SQ.SHRUB<-sqrt(RK$SHRUB)
RK$SQ.DWATCOUR<-sqrt(RK$D.WAT.COUR)
fit3<-glm(TOT.N ~ 
            OPEN.L + 
            MONT.S + 
            SQ.POLIC +
            SQ.SHRUB + 
            SQ.WATRES + 
            L.WAT.C + 
            SQ.LPROAD + 
            SQ.DWATCOUR + 
            D.PARK, 
          family=poisson, 
          data=RK)
summary(fit3)
```

Question I

The following variables have significant marginal z tests using a type 1 error rate of 0.01: Mont. S, SQ Polic, Sq Shrub, Sq Watres, L Wat C, SQ Lproad, D Park

```{r}
fit4<-glm(TOT.N ~ 
            OPEN.L + 
            MONT.S + 
            SQ.POLIC +
            SQ.SHRUB + 
            SQ.WATRES + 
            L.WAT.C + 
            SQ.LPROAD + 
            SQ.DWATCOUR + 
            D.PARK, 
          family=quasipoisson, 
          data=RK)
summary(fit4)
```

Question J

The dispersion parameter is 5.928, the parameter estimates are nearly the same for fit3 as fit4. The standard errors for fit 4 are larger than the standard errors for fit3.

Question K

```{r}
anova(fit1, fit3, test = "Chi")

```

Question L

```{r}
anova(fit2, fit4, test = "F")

```


The p value for the likelihood of ratio test is smaller than the quasilikelihood model test. In this instance I would report a p value that is too large.

```{r}
library(mgcv)
fit5 <- gam(TOT.N ~ s(D.PARK, k = 30), family = poisson, data = RK)
summary(fit5)
{
  with(RK, plot(D.PARK, TOT.N, pch = 19, main = "Road Kill Data from Zuur et al. (2009)"))
  with(RK, points(D.PARK, fitted(fit5), pch = 'x', col = 'red', type = "b")) 
}
```


```{r}
library(mgcv)
fit6 <- gam(TOT.N ~ s(D.PARK, k = 30), family = quasipoisson, data = RK)
summary(fit6)
{
  with(RK, plot(D.PARK, TOT.N, pch = 19, main = "Road Kill Data from Zuur et al. (2009)"))
  with(RK, points(D.PARK, fitted(fit6), pch = 'x', col = 'red', type = "b")) 
}
```


Question M

The GAM suggests that the data does not follow a logarithmic model until around D.Park = 10000, in addition it showcases that there is much more variance before that point than after that point indicating that a trend or a model is only viable as a predictive model when D Park > 10000 if we are using a linear model. 