---
title: "Problem Set 4"
author: 3032247297
date: "2/10/2019"
output:
  html_document:
    df_print: paged
---

### collaborated with 26642157

```{r}
library(nlme)
```


#Nested ANOVA data simulation:

Question A)

```{r}
simulate_nested <- function(group_means, sigma_beta, sigma_epsilon, q, n){
  
  p <- length(group_means)
  
  y <- 0
  group <- 0
  replicate.unique <- 0
  
  row <- 1 # Prepare for step 5
  # Step 3:
  for(i in 1:p) {
    # Step 4:
    for(j in 1:q) {
      # Step 6:
      replicate_effect <- rnorm(1, 0, sigma_beta)
      ## Increment for step 5:

        for (k in 1:n){
          group[row] <- i
          replicate.unique[row] <- j
          if (i ==1){
            replicate.unique[row] <- j
          }
          else {
            replicate.unique[row] <- j + ((i-1)*q)
          }
          y[row] <- group_means[i] + replicate_effect + rnorm(1, 0, sigma_epsilon)
          
          row <- row + 1 #iterates through row
        }
    }
  
  }
  group <- factor(group)
  replicate.unique <- factor(replicate.unique)
  
  completed <- data.frame(y, group, replicate.unique)
  completed
}
```


```{r}
set.seed(14)
# These inputs are chosen to make it easy to notice if 
# the output of this function is obviously wrong.
sim <- simulate_nested(group_means = c(100, 200, 300),
                       sigma_beta = 30,
                       sigma_epsilon = 0.5,
                       q = 3,
                       n = 2
)
sim
```


```{r}
fit <- lme(y ~ group, random = ~ 1 | replicate.unique, data = sim)
# We can look at ANOVA results like this:
anova(fit)
```


Question B)
```{r}
simulate_many_nested_p <- function(n_simulations,group_means, sigma_beta, sigma_epsilon, q, n ){
  
p_values <- 0
row <- 1

for (i in 1:n_simulations){
  sim <- simulate_nested(group_means, sigma_beta, sigma_epsilon, q, n)
  
  fit <- lme(y ~ group, random = ~ 1 | replicate.unique, data = sim)
  
  p_values[row] <- anova(fit)['group' ,'p-value'] 
  row <- row + 1
}

p_values
  
}
```


```{r}
p_sim <- simulate_many_nested_p(n_simulations = 500,
                                group_means = c(5, 10),
                                sigma_beta = 10,
                                sigma_epsilon = 20,
                                q = 10,
                                n = 3)
hist(p_sim)
```




Question C)
I think that there is far more variability in the p values that you experience when you increase the number of sub samples and have a low amount of replicates. On the contrary when you flip the two the number of p values that you experience that are statistically significant are much lower and congregate near the bottom of the plot.
```{r}
a <- simulate_many_nested_p(n_simulations = 500,
                                group_means = c(10, 20),
                                sigma_beta = 20,
                                sigma_epsilon = 10,
                                q = 2,
                                n = 25)

b <- simulate_many_nested_p(n_simulations = 500,
                                group_means = c(10, 20),
                                sigma_beta = 20,
                                sigma_epsilon = 10,
                                q = 5,
                                n = 10)

c <- simulate_many_nested_p(n_simulations = 500,
                                group_means = c(10, 20),
                                sigma_beta = 20,
                                sigma_epsilon = 10,
                                q = 10,
                                n = 5)

d <- simulate_many_nested_p(n_simulations = 500,
                                group_means = c(10, 20),
                                sigma_beta = 20,
                                sigma_epsilon = 10,
                                q = 25,
                                n = 2)

e <- simulate_many_nested_p(n_simulations = 500,
                                group_means = c(10, 20),
                                sigma_beta = 20,
                                sigma_epsilon = 10,
                                q = 50,
                                n = 1)
```

```{r}
plot(a)
plot(b)
plot(c)
plot(d)
plot(e)
```
Increasing replication increases statistical power.

uestion D)
```{r}
a <- simulate_many_nested_p(500, c(10,20), sigma_beta = 20, sigma_epsilon = 50, q = 25, n = 2)
a1 <- sum(a < 0.05) / 500

b <- simulate_many_nested_p(500, c(10,20), sigma_beta = 20, sigma_epsilon = 50, q = 25, n = 5)
b2 <- sum(b < 0.05) / 500

c <- simulate_many_nested_p(500, c(10,20), sigma_beta = 20, sigma_epsilon = 50, q = 25, n = 10)
c3 <- sum(c < 0.05) / 500

d <- simulate_many_nested_p(500, c(10,20), sigma_beta = 20, sigma_epsilon = 50, q = 25, n = 20)
d4 <- sum(d < 0.05) / 500

e <- simulate_many_nested_p(500, c(10,20), sigma_beta = 20, sigma_epsilon = 50, q = 25, n = 35)
e5 <- sum(e < 0.05) / 500

f <- simulate_many_nested_p(500, c(10,20), sigma_beta = 20, sigma_epsilon = 50, q = 25, n = 50)
f6 <- sum(f < 0.05) / 500


sub_samples <- c(2, 5, 10, 20, 35, 50)
statistical_power <- c(a1, b2, c3, d4, e5, f6)
plot(sub_samples, statistical_power, main = "number of sub-samples per replicate and statistical power")
```


Question E)
statistical power increases as you increase the number of subsamples per replicate.

Question F)
I am not sure how to do this